Portable Multi- and Many-Core Performance for Finite-Difference/Element Codes; Application to the Free-Surface Component of NEMO (NEMOLite2D 1.0).

### 2017-09-14 docs/papers/nemolite2d_performance/referee_response.txt

3/ A new section, "Related Approaches", has been added to the paper. In this
   we bring together a comparison of various other approaches including OCCA.
   This makes explicit the differences and makes clearer the advantages of
   PSyKAl.

4/ We also use the new "Related Approaches" section to emphasise that
    although the tool to generate the PSy layer exists ("PSyclone"),
    the current paper is about code structuring.

Carlos Osuna's general review:

We have added a few key lines of source code from the Momentum kernel
in section 3.3. We have also added text that explains that the
apparently poor performance of this kernel is down to the number of
division operations.  On the Ivy Bridge, division operations require
at least eight times as many cycles as a multiplication which can be
very significant.

@@

The suggestion of a "Related Work" section is a good one and we have added
this just after the Introduction.

We attribute the better performance of the SSE version of the code to the
fact that the poor SIMD efficiency of the code is failing to amortize the
larger cost of the peel and remainder loops associated with the greater
vector width of AVX. We have expanded upon this in the text.

### 2017-09-14 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: explain why SSE > AVX

Enabling SIMD vectorisation for this kernel does not significantly
improve its performance (Figure~\ref{FIG_roofline}) and in fact,
limiting it to SSE instructions (vector width of two double-precision
floating point numbers) rather than AVX (vector width of four) was
found to produce a slightly more performant version. The poor
efficiency of the SIMD version is highlighted by the fact that the
corresponding kernel performance no longer falls within the bounds of
the performance estimates produced by Habakkuk. This is because we
have incorporated the effect of SSE into that estimate by simply
assuming perfect vectorisation which gives a performance increase of a
factor of two.  Further investigation of this issue revealed that
several of the NEMOLite2D kernels make frequent use of floating-point
division operations. Although SSE and AVX versions of the division
operation are available, on Ivy Bridge they do not provide any
performance benefit~\citep{fog_instructions}. A straightforward
optimisation then would be to alter the kernels in order to reduce the
number of division operations. However, again, kernel optimisation is
outside the scope of this paper since it breaks the PSyKAl separation
of concerns.

>>>>

Enabling SIMD vectorisation for this kernel does not significantly
improve its performance (Figure~\ref{FIG_roofline}) and in fact,
limiting it to SSE instructions (vector width of two double-precision
floating point numbers) rather than AVX (vector width of four) was
found to produce a slightly more performant version. We attribute this
performance deficit to the low efficiency of the vectorised code
combined with the higher trip-counts of the peel- and remainder loops
required for the greater vector width of AVX.  The poor efficiency of
the SIMD version is highlighted by the fact that the corresponding
kernel performance no longer falls within the bounds of the
performance estimates produced by Habakkuk. This is because we have
incorporated the effect of SSE into that estimate by simply assuming
perfect vectorisation which gives a performance increase of a factor
of two.  Further investigation of this issue revealed that, as
mentioned above, several of the NEMOLite2D kernels make frequent use
of floating-point division. Although SSE and AVX versions
of the division operation are available, on Ivy Bridge they do not
provide any performance benefit~\citep{fog_instructions}. A
straightforward optimisation then would be to alter the kernels in
order to reduce the number of division operations. However, again,
kernel optimisation is outside the scope of this paper since it breaks
the PSyKAl separation of concerns.
